Pattern matching for\\ $(231,213)$-avoiding permutations

### 2015-02-02 article.tex

\begin{lemma}[Folklore]
\label{lemma:first element is 1 or n}
The 􏰅first element of any $(213,231)$-avoiding permutation
must be either $1$ or $n$.
\end{lemma}

\begin{proof} 
Any other initial element would serve as a `$2$' in either a 
$231$ or $213$ with $1$ and $n$ as the `$1$' and `$3$' respectively.
\qed
\end{proof}

@@

\begin{proposition}
\label{proposition:patternMatchingStripe}
Let $\ptext$ and $\pmotif$ be two permutations avoiding (213,231).
$\pmotif$ is a motif of $\ptext$ if and only if the 
stripe of $\pmotif$ is a motif of the stripe of $\ptext$.
\end{proposition}

@@

%We need the following lemmata
%(the proofs of Lemma~\ref{lemma:upup} and Lemma~\ref{lemma:downdown} 
%are analogous and we omit them).
%
%\begin{lemma}
%	\label{lemma:updown}
%	Let $\pi \in \AV_n(213,231)$.
%	For all $1 \leq i < j \leq n$, 
%	if $\stripe{\pi}{i}=\ustep$
%	and $\stripe{\pi}{j}=\dstep$
%	then $\pi[i] < \pi[j]$.
%\end{lemma}
%		
%\begin{proof} 
%By induction on the size of the permutation.
%The proposition is certainly true for the $12$ and $21$.
%Suppose the proposition holds true for $\AV_n(213,231)$.
%Let $\pi \in \AV_{n+1}(213,231)$ and
%$\pi'$ be the permutation $\pi[2:]$.
%$\pi' \in \AV_n(213,231)$.
%By induction hypothesis, we only need to compare the first
%element of $\pi'$ to every element in $\pi'$.
%According to lemma~\ref{lemma:first element is 1 or n},
%the first element of $\pi$ is either $1$ or $n$.
%If the first element of $\pi$ is $1$ then $s_\pi(1) = a$
%and $\pi(i) < \pi(j)$ for every $2 \leq j \leq n$.
%If the first element of $\pi$ is $n$ then
% $s_\pi(1) = d$.
%\qed
%\end{proof}
%
%\begin{lemma}
%	\label{lemma:upup}
%	Let $\pi \in \AV_n(213,231)$.
%	For all $1 \leq i < j \leq n$,
%	if $\stripe{\pi}{i}=\ustep$
%	and $\stripe{\pi}{j}=\ustep$
%	then $\pi[i] < \pi[j]$.
%\end{lemma}
%		
%
%\begin{lemma}
%	\label{lemma:downdown}
%	Let $\pi \in \AV_n(213,231)$.
%	For all $1 \leq i < j \leq n$,
%	if $\stripe{\pi}{i} = \dstep$
%	and $\stripe{\pi}{j} = \dstep$
%	then $\pi[i] > \pi[j]$.
%\end{lemma}
%
%
%Most of the interest of the above lemmata stems from the following proposition.
%
%\begin{proposition}
%	\label{proposition:patternMatchingStripe}
%	Let $\ptext$ and $\pmotif$ be two permutations avoiding (213,231).
%	$\pmotif$ is a motif of $\ptext$ if and only if the 
%	stripe of $\pmotif$ is a motif of the stripe of $\ptext$.
%\end{proposition}
%
%		
%		\textbf{Proof} \\
%		$\Rightarrow$ Suppose that $\pmotif$ is a motif of $\ptext$
%		by subsequence $s$.
%		By definition $s$ normalized is equal to
%		$\pmotif$ so they have the same stripe.\\ 
%		$\Leftarrow$ 
%		Let $\ptext$ be a permutation avoiding $(213,231)$, 
%		$S_1$ be its stripe, 
%		$\pmotif$  be a permutation avoiding $(213,231)$, 
%		and $S_2$ its stripe.
%		Let $\phi$ be the function
%		that match $S_2$ to $S_1$.
%		We define $\varphi$ as, 
%		$\forall x \in [|\pmotif|-1]$,
%		$\varphi(x) = \phi(x)$,
%		and $\varphi(|\pmotif|) = \phi(|\pmotif|-1)+1$
%		Suppose that $\exists x,y \in [|\pmotif|]$, 
%		$\pmotif(x)<\pmotif(y)$ 
%		and $\ptext(\varphi(x))>\ptext(\varphi(y))$,
%		without lost of generality we can suppose that $x<y$,
%		Let see all the case possible that can make this 
%		hypothesis possible :
%		\begin{itemize}		
%			\item if $\pmotif(x)<\pmotif(y)$, 
%			$\stripe{\pmotif}{x}=\ustep$ 
%			and $\stripe{\pmotif}{y}=\ustep$
%			(if x and y are increasing in $\pmotif$),
%			then $\stripe{\ptext}{\varphi(x)} =\ustep$ and $\stripe{\ptext}{\varphi(y)}=\ustep$,
%			(the match are also increasing).
%			As $x<y$ and as $\varphi$ is an increasing function, 
%			$\varphi(x) < \varphi(y)$.
%			And thus, by \textbf{Lemma \ref{lemma:upup}},
%			$\ptext(\varphi(x)) < \ptext(\varphi(y))$ .
%			Which contradict our hypothesis,
%			so this case is not possible.
%		
%			\item if $\pmotif(x)<\pmotif(y)$, 
%			$\stripe{\pmotif}{x}=\ustep$ 
%			and $\stripe{\pmotif}{y}=\dstep$
%			then $\stripe{\ptext}{\varphi(x)}=\ustep$ and $\stripe{\ptext}{\varphi(y)}=\dstep$,
%			As $x<y$ and as $\varphi$ is an increasing function, 
%			$\varphi(x) < \varphi(y)$.
%			And thus, by \textbf{Lemma \ref{lemma:updown}},
%			$\ptext(\varphi(x)) < \ptext(\varphi(y))$ .
%			Which contradict our hypothesis,
%			so this case is not possible.			
%
%			\item if $\pmotif(x)>\pmotif(y)$, 
%			$\stripe{\pmotif}{x}=\dstep$ 
%			and $\stripe{\pmotif}{y}=\dstep$
%			then $\stripe{\ptext}{\varphi(x)}=\dstep$ and $\stripe{\ptext}{\varphi(y)}=\dstep$,
%			As $x<y$ and as $\varphi$ is an increasing function, 
%			$\varphi(y) < \varphi(x)$.
%			And thus, by \textbf{Lemma \ref{lemma:downdown}},
%			$\ptext(\varphi(y)) > \ptext(\varphi(x))$ .
%			Which contradict our hypothesis,
%			so this case is not possible.
%			
%			\item if $\pmotif(x)>\pmotif(y)$, 
%			$\stripe{\pmotif}{x}=\dstep$ 
%			and $\stripe{\pmotif}{y}=\ustep$
%			then $\stripe{\ptext}{\varphi(x)}=\dstep$ and $\stripe{\ptext}{\varphi(y)}=\ustep$,
%			As $x<y$ and as $\varphi$ is an increasing function, 
%			$\varphi(y) < \varphi(x)$.
%			And thus, by \textbf{Lemma \ref{lemma:updown}},
%			$\ptext(\varphi(y)) > \ptext(\varphi(x))$ .
%			Which contradict our hypothesis,
%			so this case is not possible.$\Box$
%			
%			\item If $y = |\pmotif|$ and $x=|\pmotif|-1$,
%			Clearly if $\stripe{\pmotif}{x}=\ustep$,
%			then $\pmotif(y)-\pmotif(x)>0$ 
%			and because we have a match in the ribbon
%			$\ptext(\varphi(y))-\ptext(\varphi(x))>0$.
%			Same idea goes if $\stripe{\pmotif}{x}=\dstep$.
%			So this case is not possible.
%			
%			\item if $y = |\pmotif|$ and $x<|\pmotif|-1$,
%			if $\stripe{\pmotif}{x}=\ustep$ ie
%			$\pmotif(x+1)-\pmotif(x)>0$. Then
%			we must have either
%			$\pmotif(x)<\pmotif(x+1)<\pmotif(|\pmotif|)$
%			or $\pmotif(x)<\pmotif(|\pmotif|)<\pmotif(x+1)$ 
%			to avoid $231$ or $213$ in the motif.
%			Because we have a match in the stripe
%			and what we prove above,
%			$\ptext(\varphi(x+1)-\ptext(\varphi(x))>0$.
%			But by hypothesis $\ptext(\varphi(|\ptext|))<\ptext(\varphi(x))$,
%			so we have either 		
%			$\pmotif(|\pmotif|)<\pmotif(x)<\pmotif(x+1)$
%			or $\pmotif(|\pmotif|)<\pmotif(x)<\pmotif(x+1)$ 
%			which both lead to a $231$ or a $213$ motif
%			in the text which is not possible.
%			The same idea goes if $\stripe{\pmotif}{x}=\dstep$. 			
%			
%		\end{itemize}
%		So our hypothesis is not realisable.So our assertion is true. $\Box$\\

>>>>

\begin{lemma}[Folklore]
\label{lemma:first element is 1 or n}
The first element of any $(213,231)$-avoiding permutation
must be either $1$ or $n$.
\end{lemma}

\begin{proof} 
Any other initial element would serve as a `$2$' in either a 
$231$ or $213$ with $1$ and $n$ as the `$1$' and `$3$' respectively.
\end{proof}

@@

\begin{proposition}
Let $\sigma \in \AV_k(213,231)$ and $\pi \in \AV_k(213,231)$.
$\sigma$ is a motif of $\pi$ if and only if 
$s_\sigma$ is a subsequence of $s_\pi$.
\end{proposition}

		% without lost of generality we can suppose that $x<y$,
% 		Let see all the case possible that can make this 
% 		hypothesis possible :
% 		\begin{itemize}		
% 			\item if $\pmotif(x)<\pmotif(y)$, 
% 			$\stripe{\pmotif}{x}=\ustep$ 
% 			and $\stripe{\pmotif}{y}=\ustep$
% 			(if x and y are increasing in $\pmotif$),
% 			then $\stripe{\ptext}{\varphi(x)} =\ustep$ and $\stripe{\ptext}{\varphi(y)}=\ustep$,
% 			(the match are also increasing).
% 			As $x<y$ and as $\varphi$ is an increasing function, 
% 			$\varphi(x) < \varphi(y)$.
% 			And thus, by \textbf{Lemma \ref{lemma:upup}},
% 			$\ptext(\varphi(x)) < \ptext(\varphi(y))$ .
% 			Which contradict our hypothesis,
% 			so this case is not possible.

@@
